### 2022-10-11 AISTATS2023/paper.tex
with Eirini's review until chapter 1

Recently, Machine Learning (ML) has been adopted in critical domains,
such as healthcare and finance. In these areas, we need a combination
of accurate predictions along with meaningful explanations to support
them. For this reason, there is an increased interest in Explainable
AI (XAI) to understand the decision mechanism of complex black-box
models. XAI literature distinguishes between local and global
techniques~\citep{Molnar2020interpretable}. Local methods provide
instance-level explanations, i.e.~explain the prediction for a
specific instance, whereas global methods summarize the entire model
behavior.

Global methods create a global explanation by aggregating the
instance-level explanations into a single interpretable outcome,
usually a number or a plot. When strong interactions between features
exist, the instance-level explanations may significantly deviate from
the aggregated outcome~\citep{herbinger2022repid}, a phenomenon called
\emph{aggregation bias}~\citep{mehrabi2021survey}. Therefore, global
methods should quantify not only the aggregated (averaged) effect, but
also the level of heterogeneity of instance-level explanations. The
same holds for Feature Effect (FE)~\citep{Gromping2020MAEP}, a class
of global XAI methods that isolate the impact of a single feature on
the output.\footnote{FE methods also isolate the combined effect of a
  pair of features to the output. Combinations of more than two
  features are uncommon, since they are difficult to estimate and
  visualize.} For example, a feature with zero average effect may
indicate no effect on the output or a highly positive effect in some
cases and a highly negative effect in others. The ambiguity about the
correct interpretation can be resolved, if the FE method informs about
the heterogeneity of instance-level explanations, a quantity we refer
to as the \emph{uncertainty of the effect}.

There are three widely-used FE methods: \emph{Partial Dependence
  Plots} (PDP)\citep{friedman2001greedy}, \emph{Marginal Plots}
(MP)\citep{apley2020visualizing} and \emph{Aggregated Local Effects}
(ALE)\citep{apley2020visualizing}. ALE has been established as the
state-of-the-art for measuring the average effect, since PDP and MP
have been criticized~\citep{Gromping2020MAEP} to compute erroneous
effects when the input features are (highly) correlated. However, ALE
faces two limitations, the first concerns ALE definition and the
second ALE approximation. Regarding the definition, ALE does not
inform the user about the uncertainty. In contrast, in PDP the
heterogeneity of instance-level effects can be identified by exploring
the Individual Conditional Expectations
(ICE)~\citep{goldstein2015peeking}. Regarding the approximation, ALE
requires partitioning the feature domain into non-overlapping
intervals (bin-splitting) and estimate the effect inside each interval
(bin-effect) from the training set instances that fall inside. Finding
an appropriate sequence of bins is of particular importance, since
ALE's interpretation is meaningful only inside each
bin~\citep{molnar2022}. However, this crucial step has not raised the
appropriate attention. With the current
approaches~\citep{apley2020visualizing, gkolemis22}, the domain is
split \emph{blindly}, i.e.~without considering the characteristics of
the instance-level effects, in \(K\) equisized bins.

In this paper, we present Uncertainty-aware Accumulated Local Effects
(UALE), a method that visualizes in a single plot both the average
feature effect and the uncertainty, i.e.~the level of heterogeneity of
instance-level effects. The average effect is defined as in ALE and
the uncertainty inherits ALE's advantages, i.e.~takes into account the
correlations between features. We also propose an accurate UALE
approximation. We formally prove that for an unbiased estimation of
the uncertainty, the partitioning of the feature domain must respect
specific conditions. Using that, we set-up an optimization problem
where we search for a partitioning, a sequence of \emph{variable-size}
intervals, (a) with enough instances (accurate estimation) and (b)
that respect the above conditions (unbiased estimation). Finally, we
propose a method for automatically finding the optimal
partitioning. We apply our method in multiple synthetic and real world
datasets for evaluating its performance. The contributions of this
paper are:

\begin{itemize}
\item UALE, a feature effect method that quantifies both the average
  effect and the uncertainty, i.e.,~the heterogeneity of instance level
  effects.
\item A formal proof of the conditions the feature domain partitioning
  must fulfill for an unbiased approximation of the uncertainty.
\item An accurate UALE approximation by automatically partitioning
  into variable-size intervals that respect the above conditions.
\end{itemize}

@@

\section{BACKGROUND AND RELATED WORK}

We here review the existing methods for modeling the average effect
and the heterogeneity of instance-level effects and we focus on ALE
definition~\citep{apley2020visualizing} and
approximation~\citep{apley2020visualizing, gkolemis22}.

\paragraph{Notation.} We refer to random variables (rv) using
uppercase \( X \), to simple variables with plain lowercase \( x \)
and to vectors with bold \( \xb \). Often, we partition the input
vector \(\xb \in \mathbb{R}^D\) to the feature of interest
\(x_s \in \mathbb{R} \) and the rest of the features
\(\xc \in \mathbb{R}^{D-1}\). For convenience we denote it as
\((x_s, \mathbf{x}_c)\), but we it refers the vector
\((x_1, \cdots , x_s, \cdots, x_D)\). Equivalently, we denote the
corresponding rv as \(\mathbf{X} = (X_s, \mathbf{X}_c)\). When we
refer to an instance of the training set, we use
\(\xb^i= (x_s^i, \xc^i) \).

% The black-box function is
% \(f : \mathbb{R}^D \rightarrow \mathbb{R}\) and the FE of the \(s\)-th
% feature is \(f^{\mathtt{<method>}}(x_s)\), with \(\mathtt{<method>}\)
% indicating the particular method in use.\footnote{An extensive list of
%   all symbols used in the paper is provided at the Appendix.}

\subsection{Feature Effect Methods}
\label{sec:feat-effect-meth}

The three well-known feature effect methods are: PDP, MP and ALE. PDP
formulates the FE of the \(s\)-th feature as an expectation over the
marginal distribution \(\mathbf{X}_c\), i.e.,
\(f^{\mathtt{PDP}}(x_s) =
\mathbb{E}_{\mathbf{X}_c}[f(x_s,\mathbf{X}_c)]\), whereas MP
formulates it as an expectation over the conditional
\(\mathbf{X}_c|X_s\), i.e.,
\(f^{\mathtt{MP}}(x_s) = \mathbb{E}_{\mathbf{X}_c|x_s}[f(x_s,
\mathbf{X}_c)]\). ALE defines the local effect at \(x_s = z\) as
\(f^s(z, \xc) = \frac{\partial f}{\partial x_s} (z, \xc)\). The local
explanations are then weighted by the conditional distribution
\(p(\xc|x_s=z)\) to compute the averaged effect \(\mu(z)\). ALE is the
accumulation of the averaged local effects:

@@

Eq.~(\ref{eq:ALE}) has important advantages compared to PDP and MP in
cases of correlated features. In these cases, PDP integrates over
unrealistic instances due to the use of the marginal distribution
\(\mathbf{X}_c \), and MP computes aggregated effects, i.e., imputes
the combined effect of sets of features to a single feature. ALE
resolves both issues.

\subsection{Heterogeneity Of Instance-Level Effects.}
\label{sec:quant-heter-effects}

FE methods show what is the average (global) effect on the output, if
the value of a specific feature is increased/decreased. For completing
the explanation about a feature, we need to also know to what extent
the local effects deviate from the global explanation, i.e.~to
quantify the uncertainty of the average effect.

Heterogeneity It comes naturally to also ask \textit{to
  what extent }. Quantifying the uncertainty of the global explanation
has attracted a lot of interest. ICE and
d-ICE\citep{goldstein2015peeking} provide a set of curves that are
plot on top-of the PDP. Both methods produce one curve for each
instance of the training set;
\(f^{\mathtt{ICE}}_i(x_s) = f(x_s, \xc^i)\) for ICE and
\(f^{\mathtt{d-ICE}}_i(x_s) = \frac{\partial f}{\partial x_s} (x_s,
\xc^i)\) for d-ICE. The user then visually observes if the curves are
homogeneous and to what extent they deviate from the PDP. Some methods
try to automate the aforementioned visual exploration, by grouping
(d-)ICE plots into clusters~\citep{molnar2020model,
  herbinger2022repid, britton2019vine}. Some other approaches, like
H-Statistic\citep{friedman2008predictive}, Greenwell's interaction
index\citep{greenwell2018simple} or SHAP interaction
values\citep{lundberg2018consistent}, quantify the level of
interaction between the input features, with an interaction value. A
strong interaction index is an indicator for the existence of
heterogeneous effects.

The aforementioned approaches are under two limitations; They either
do not quantify the uncertainty of the FE directly or they are based
on PDPs, and, therefore, they are subject to the failure modes of PDPs
in cases of correlated features\citep{baniecki2021fooling}, as we will
show in-depth in Section~\ref{sec:simulation-examples-1}. To the best
of our knowledge, no work exist so far that quantifies the
heterogeneous effects based on the formulation of ALE.

@@

In ML scenarios, ALE is estimated from the training set
examples. Therefore, \citep{apley2020visualizing} proposed dividing
the feature domain in \(K\) equal-sized bins and estimating the local effects in
each bin by evaluating the black box-function at the bin limits:
\begin{equation}
  \label{eq:ALE_accumulated_mean_est}
  \hat{f}^{\mathtt{ALE}}(x_s) = \sum_{k=1}^{k_x} \frac{1}{|\mathcal{S}_k|} \sum_{i:\mathbf{x}^i \in
    \mathcal{S}_k} \left [ f(z_{k}, \xc^i) - f(z_{k-1}, \xc^i)) \right ]
\end{equation}
We denote as \(k_x\) the index of the bin that \(x_s\) belongs to,
i.e. \(k_x: z_{k_x-1} \leq x_s < z_{k_x} \) and \(\mathcal{S}_k\) is
the set of training instance that lie in the \(k\)-th bin, i.e.
\( \mathcal{S}_k = \{ \xb^i : z_{k-1} \leq x^i_s < z_{k} \}
\). Afterwards,~\citep{gkolemis22} proposed the Differential ALE
(DALE) that computes the local effects on the training instances using
auto-differentiation:
\begin{equation}
  \label{eq:DALE_accumulated_mean_est}
  \hat{f}^{\mathtt{DALE}}(x_s) = \Delta x \sum_{k=1}^{k_x} \frac{1}{|\mathcal{S}_k|} \sum_{i:\mathbf{x}^i \in
    \mathcal{S}_k} f^s(\mathbf{x}^i)
\end{equation}
%
Their method has the advantages of remaining on-distribution even when
bins become wider and, most importantly, allows the recomputation of
the accumulated effect with different bin-splitting with near-zero
computational cost.

Both approximations ask from the user to blindly decide the number of
bins, denoted with \(K\), for splitting the axis into equally-sized
bins. This approach introduces significant limitations. Specifically, setting \(K\) to a small
value may hide fine-grain effects due to large bins while setting \(K\)
to a high value leads to poor bin-effect estimations from limited
samples. In general, the user may face contradictory explanations for
different \(K\) without a clue for which one to trust.

@@

As before, no interaction exits \(\alpha=0\), therefore, the
ground-truth uncertainty is zero, i.e.
\(f^{\mathcal{GT}}_{\sigma^2}(x_1) = 0\). The weigts are
\(a_1=2, a_2=0.5\), therefore, the ground-truth effect is
\(f_\mu^{\mathtt{GT}}(x_1)\) is, \(2x_1\) when \(0 \leq x_1 < 0.2\),
\(-2x_1\) when \(0.2 \leq x_1 < 0.4\) and zero otherwise. In
Figure~\ref{fig:ex-synth-1-2}, we observe that PDP estimation is
completely opposite to the ground-truth effect, i.e.~negative in the
region \([0, 0.2)\) and positive in \([0.2, 0.4)\), and the ICE
erroneously implies the existence of heterogeneous effects. As before,
ALE quantifies correctly the ground truth effect, the
zero-uncertainty and extracts the constant effect regions.

>>>>

Recently, Machine Learning (ML) has been adopted in critical domains,
such as healthcare and finance. In these areas, we need a combination
of accurate predictions along with meaningful explanations to support
them. For this reason there is an increased interest in Explainable AI
(XAI), to understand the decision mechanism of complex black-box
models. XAI literature distinguishes between local and global
techniques~\citep{Molnar2020interpretable}. Local methods provide
instance-level explanations, i.e.~explain the prediction for a
specific instance, whereas global methods summarize the entire model
behavior.

Global methods create a global explanation by aggregating the
instance-level explanations into a single interpretable outcome,
usually a number or a plot.
A popular class of global methods are the so-called Feature Effect (FE) methods~\citep{Gromping2020MAEP} that isolate the impact of a single feature on
the output\footnote{FE methods also isolate the combined effect of a
  pair of features to the output. Combinations of more than two
  features are uncommon, since they are difficult to estimate and
  visualize.}. There are three widely-used FE methods: \emph{Partial Dependence
  Plots} (PDP)\citep{friedman2001greedy}, \emph{Marginal Plots}
(MP)\citep{apley2020visualizing} and \emph{Aggregated Local Effects}
(ALE)\citep{apley2020visualizing} with ALE being established as the
state-of-the-art for measuring the average effect, since PDP and MP
have been criticized~\citep{Gromping2020MAEP} to compute erroneous
effects when the input features are (highly) correlated.

However,
when strong interactions between features exist,
the instance-level feature effects may significantly deviate from
the aggregated outcome of ALE~\citep{herbinger2022repid}, a phenomenon called
\emph{aggregation bias}~\citep{mehrabi2021survey}. 
Aggregation bias leads to ambiguities; for example, a feature with zero average effect may
indicate no effect on the output or a highly positive effect for some instances and a highly negative effect for others. Such ambiguities could be resolved if not only the average effect but also the deviation of instance-level feature effects, a quantity we refer to as the \emph{uncertainty of the effect}, from the average effect is provided.

%Therefore, global methods should quantify not only the aggregated (averaged) effect, but also the level of heterogeneity of instance-level explanations.

%There are three widely-used FE methods: \emph{Partial Dependence Plots} (PDP)\citep{friedman2001greedy}, \emph{Marginal Plots}
%(MP)\citep{apley2020visualizing} and \emph{Aggregated Local Effects}
%(ALE)\citep{apley2020visualizing}. ALE has been established as the
%state-of-the-art for measuring the average effect, since PDP and MP
%have been criticized~\citep{Gromping2020MAEP} to compute erroneous
%effects when the input features are (highly) correlated.

In this work, we propose UALE,  a global feature-effect method that extends ALE by quantifying the uncertainty associated with the average effect. 
Similarly to ALE, UALE's approximation requires partitioning the feature domain into non-overlapping intervals (bin-splitting). In contrast to ALE that requires a fixed equi-width partitioning without considering the characteristics of
the instance-level effects, we propose an automatic partitioning that minimizes the uncertainty of the resulting bins.

%However, ALE faces two limitations, the first concerns ALE definition and the second ALE approximation. Regarding the definition, ALE does not inform the user about the uncertainty. 
\color{blue}\todo{This should go to related work}In contrast, in PDP the
heterogeneity of instance-level effects can be identified by exploring
the Individual Conditional Expectations
(ICE)~\citep{goldstein2015peeking}. \color{black}
%Regarding the approximation, ALE requires partitioning the feature domain into non-overlapping intervals (bin-splitting) and estimate the effect inside each interval (bin-effect) from the training set instances that fall inside. 
\color{blue}\todo{to be moved in related work}
Finding
an appropriate sequence of bins is of particular importance, since
ALE's interpretation is meaningful only inside each
bin~\citep{molnar2022}. However, this crucial step has not raised the
appropriate attention. With the current
approaches~\citep{apley2020visualizing, gkolemis22}, the domain is
split \emph{blindly}, i.e.~without considering the characteristics of
the instance-level effects, in \(K\) equisized bins.\color{black}

%In this paper, we present Uncertainty-aware Accumulated Local Effects(UALE), a method that visualizes in a single plot both the av erage feature effect and the uncertainty, i.e.~the level of heterogeneity of instance-level effects. The average effect is defined as in ALE and the uncertainty inherits ALE's advantages, i.e.~takes into account the correlations between features. We also propose an accurate UALE approximation. We formally prove that for an unbiased estimation of the uncertainty, the partitioning of the feature domain must respect specific conditions. Using that, we set-up an optimization problem where we search for a partitioning, a sequence of \emph{variable-size}intervals, (a) with enough instances (accurate estimation) and (b) that respect the above conditions (unbiased estimation). Finally, we propose a method for automatically finding the optimal partitioning. We apply our method in multiple synthetic and real world datasets for evaluating its performance. 
The contributions of this
paper are:
\begin{itemize}
\item A global feature effect method that quantifies both the average
  effect and the uncertainty, i.e.,~the heterogeneity of instance level
  effects.
\item An unbasied approximation of the uncertainty through a dynamic uncertainty-driven bin partitioning of the feature domain.
%A formal proof of the conditions the feature domain partitioning must fulfill for an unbiased approximation of the uncertainty.
\item A thorough evaluation on both synthetic and real datasets.
%An accurate UALE approximation by automatically partitioning into variable-size intervals that respect the above conditions.
\end{itemize}

@@

\section{BACKGROUND AND RELATED WORK}
The necessary background for ALE is presented in Section~\ref{sec:feat-effect-meth}), whereas ALE approximations are presented in Section~\ref{sec:ale-approximation}.

The approximation 
In this section, we describe the basic methods for FE and uncertainty
quantification, focusing on ALE definition. We, then, review the ALE
approximation\citep{apley2020visualizing, gkolemis22}, describing some
of its vulnerabilities.

\paragraph{Notation}

Let \(\mathcal{X} \in \mathbb{R}^d\) be the d-dimensional feature space, \(\mathcal{Y} \in \mathbb{R}\) the target space and \(f(\cdot) : \mathcal{X} \rightarrow \mathcal{Y}\) the black-box function. We use index \(s \in \{1, \ldots, d\}\) for the feature of interest and \(c \in \{1, \ldots, d\} - s\) for the rest feature indexes. For convenience, we denote the feature vector \(\xb = (x_1, \cdots , x_s, \cdots, x_D)\) with \((x_s, \xc)\) and the corresponding random variables \(X = (X_1, \cdots , X_s, \cdots, X_D)\) with \((X_s, X_c)\). The training set is \(\mathcal{D} = \{(\xb^i, y^i)\}_{i=1}^N\) sampled i.i.d. from the distribution \(\mathbb{P}_{X,Y}\). Finally, we use \(f^{\mathtt{<method>}}(x_s)\) for the FE of the \(s\)-th feature, with \(\mathtt{<method>}\)
indicating the particular method in use, for example ALE.\footnote{An extensive list of all
  symbols used in the paper is provided at the Appendix.}

\subsection{Feature Effect Methods and ALE}
\label{sec:feat-effect-meth}
The three well-known feature effect methods are: \emph{Partial Dependence
  Plots} (PDP)\citep{friedman2001greedy}, \emph{Marginal Plots}
(MP)\citep{apley2020visualizing} and \emph{Aggregated Local Effects}
(ALE)\citep{apley2020visualizing}
PDP
formulates the FE of the \(s\)-th attribute as an expectation over the
marginal distribution \(\mathbf{X}_c\), i.e.,
\(f^{\mathtt{PDP}}(x_s) =
\mathbb{E}_{\mathbf{X}_c}[f(x_s,\mathbf{X}_c)]\), whereas MP
formulates it as an expectation over the conditional
\(\mathbf{X}_c|X_s\), i.e.,
\(f^{\mathtt{MP}}(x_s) = \mathbb{E}_{\mathbf{X}_c|x_s}[f(x_s,
\mathbf{X}_c)]\). ALE defines the local effect of the \(s\)-th feature
at instance \(x_s = z\) as
\(f^s(z, \xc) = \frac{\partial f}{\partial x_s} (z, \xc)\). All the
local explanations at \(z\) are, then, weighted by the conditional
distribution \(p(\xc|x_s=z)\) and are averaged, to produce the
averaged effect \(\mu(z)\). ALE is the accumulation of the averaged
local effects:

@@

Eq.~(\ref{eq:ALE}) has specific advantages which gain particular value in cases of
correlated input features. In these cases, PDP integrates over
unrealistic instances due to the use of the marginal distribution
\(\mathbf{X}_c \), and MP computes aggregated effects, i.e., imputes
the combined effect of sets of features to a single feature. ALE
manages to resolve both issues, and is therefore the most trustable
method in cases of correlated features. \todo{cite someone}

\subsection{Quantification Of Heterogeneous Effects.}
\label{sec:quant-heter-effects}

FE methods reply to the question \textit{what is the average (global)
  effect on the output, if the value of a specific feature is
  increased/decreased}. It comes naturally to also ask \textit{to what
  extent the local effects \todo{add what local means here} deviate from the global (average)
  effect/explanation}. 
  \todo{A connector line is needed here on the connection between local-global and uncertainty}
  Quantifying the uncertainty of the global explanation
has attracted a lot of interest. ICE and
d-ICE\citep{goldstein2015peeking} provide a set of curves that are
plot on top-of the PDP. Both methods produce one curve for each
instance of the training set;
\(f^{\mathtt{ICE}}_i(x_s) = f(x_s, \xc^i)\) for ICE and
\(f^{\mathtt{d-ICE}}_i(x_s) = \frac{\partial f}{\partial x_s} (x_s,
\xc^i)\) for d-ICE. The user then visually observes if the curves are
homogeneous, i.e., all instances have similar values, and to what extent they deviate from the PDP. Some methods
try to automate the aforementioned visual exploration, by grouping
(d-)ICE plots into clusters~\citep{molnar2020model,
  herbinger2022repid, britton2019vine}. Some other approaches, like
H-Statistic\citep{friedman2008predictive}, Greenwell's interaction
index\citep{greenwell2018simple} or SHAP interaction
values\citep{lundberg2018consistent}, quantify the level of
interaction between the input features, with an interaction value \todo{not clear what htis value is, maybe add a few words}. A
strong interaction index is an indicator for the existence of
heterogeneous effects.

The aforementioned approaches are under two limitations; They either
do not quantify the uncertainty of the FE directly \todo{I dont get this argument. What do you mean directly?} or they are based
on PDPs, and, therefore, they are subject to the failure modes of PDPs
in cases of correlated features\citep{baniecki2021fooling}, as we will also show in our experimental analysis (c.f., Section~\ref{sec:simulation-examples-1}). To the best
of our knowledge, no work exist so far that quantifies the
heterogeneous effects based on the formulation of ALE.
\todo{My understanding of the contribution is also the last sentence. None has done it for ALE, but it has been done f for PDPs. For this I think we need some more related owrk on the limiations of PDPs. Also, is this work a straight-forward transfer of the d-ICE etc methods decribed above or transfering these ideas to ALE requires a different approach?}

@@

In ML scenarios, ALE is estimated from the training set
examples. Therefore, \citep{apley2020visualizing} proposed dividing
the feature domain in \(K\) equal-sized bins and estimating the local effects in
each bin by evaluating the black box-function $f()$ at the bin limits: \todo{k or K? I prefer k for the cardinality and it seems that k is used afterwards. }
\begin{equation}
  \label{eq:ALE_accumulated_mean_est}
  \hat{f}^{\mathtt{ALE}}(x_s) = \sum_{k=1}^{k_x} \frac{1}{|\mathcal{S}_k|} \sum_{i:\mathbf{x}^i \in
    \mathcal{S}_k} \left [ f(z_{k}, \xc^i) - f(z_{k-1}, \xc^i)) \right ]
\end{equation}
We denote as \(k_x\) the index of the bin that \(x_s\) belongs to,
i.e. \(k_x: z_{k_x-1} \leq x_s < z_{k_x} \) and \(\mathcal{S}_k\) is
the set of training instance that lie in the \(k\)-th bin, i.e.
\( \mathcal{S}_k = \{ \xb^i : z_{k-1} \leq x^i_s < z_{k} \}
\).

\citep{gkolemis22} proposed the Differential ALE
(DALE) that computes the local effects on the training instances using
auto-differentiation:
\begin{equation}  \label{eq:DALE_accumulated_mean_est}
  \hat{f}^{\mathtt{DALE}}(x_s) = \Delta x \sum_{k=1}^{k_x} \frac{1}{|\mathcal{S}_k|} \sum_{i:\mathbf{x}^i \in
    \mathcal{S}_k} f^s(\mathbf{x}^i)
\end{equation}
%
Their method has the advantages of remaining on-distribution even when
bins become wider and, most importantly, allows the recomputation of
the accumulated effect with different bin-splitting with near-zero additional
computational cost.

Both approximations, however, require the number of bins \(K\) as input and perform an  equally-sized bin splitting of the attribute domain. The choice of \(K\) is therefore critical and introduces significant limitations. Specifically, 
setting \(K\) to a small
value may hide fine-grain effects due to large bins while setting \(K\)
to a high value leads to poor bin-effect estimations from limited
samples. In general, the user may face contradictory explanations for
different \(K\) values without a clue for which one to trust.

@@

As before, no interaction exits \(\alpha=0\), therefore, the
ground-truth uncertainty is zero, i.e.
\(f^{\mathcal{GT}}_{\sigma^2}(x_1) = 0\). The weights are
\(a_1=2, a_2=0.5\), therefore, the ground-truth effect is
\(f_\mu^{\mathtt{GT}}(x_1)\) is, \(2x_1\) when \(0 \leq x_1 < 0.2\),
\(-2x_1\) when \(0.2 \leq x_1 < 0.4\) and zero otherwise. In
Figure~\ref{fig:ex-synth-1-2}, we observe that PDP estimation is
completely opposite to the ground-truth effect, i.e.~negative in the
region \([0, 0.2)\) and positive in \([0.2, 0.4)\), and the ICE
erroneously implies the existence of heterogeneous effects. As before,
ALE quantifies correctly the ground truth effect, the
zero-uncertainty and extracts the constant effect regions.
